---
title: "RockTree"
author: "Matt Russell"
date: "11/16/2020"
output: html_document
---
Load libraries:

```{r echo = F, warning = FALSE, message = FALSE}
library(tidyverse)
library(maps)
library(geosphere)
library(ggrepel)
```

Load data:

```{r echo = F, warning = FALSE, message = FALSE}
tree <- read_csv('rocktree.csv')
```

Sort by number of species:

```{r echo = F, warning = FALSE, message = FALSE}
tree %>% 
  group_by(Species) %>% 
  summarize(num = n()) %>% 
  arrange(desc(num))
```

Graph height of tree over time:

```{r echo = F, warning = FALSE, message = FALSE}
tree_2000 <- tree %>% 
  filter(Year >= 2000) %>% 
  summarize(mean_ht = mean(Height_ft))
            
ggplot(tree, aes(Year, Height_ft)) +
  geom_point() +
  stat_smooth(method = "lm") +
  labs(y = "Height of Rockefeller tree (feet)") +
  theme_bw()
```

Graph mean height of tree by state: 

```{r echo = F, warning = FALSE, message = FALSE}
tree %>% 
  group_by(OriginState) %>% 
  summarize(num = n(),
            mean_ht = mean(Height_ft, na.rm = T)) %>% 
  arrange(desc(mean_ht)) %>% 
  filter(num >2 & !is.na(OriginState)) %>%
  ggplot(aes(x = reorder(OriginState, mean_ht), y = mean_ht)) +
  geom_bar(stat = "identity", fill = "lightgreen", col = "black") +
  coord_flip()+
  labs(y = "Mean height of Rockefeller tree (feet)", x = "State") +
  theme_bw()
```

Make state and county data sets:

```{r}
states <- map_data("state")

tree_states <- subset(states, region %in% c("connecticut", "pennsylvania", "new york", 
                                            "ohio", "new jersey", "vermont", 
                                            "massachusetts","new hampshire"))
counties <- map_data("county")
tree_counties <- subset(counties,  region %in% c("connecticut", "pennsylvania", "new york", 
                                            "ohio", "new jersey", "vermont", 
                                            "massachusetts", "new hampshire"))
```

Map states:

```{r}
state_map <- ggplot(data = tree_states, mapping = aes(x = long, y = lat, group = group)) + 
  geom_polygon(color = "black", fill = "white") +
  coord_map("conic", lat0 = 40) +
  theme_bw()
state_map
```

Map counties:

```{r}
county_map <- ggplot(data = tree_states, mapping = aes(x = long, y = lat, group = group)) + 
  geom_polygon(data = tree_counties, fill = NA, color = "gray") +
  geom_polygon(color = "black", fill = NA) +
  coord_map("conic", lat0 = 40) +
  theme_bw()
county_map
```

Map states with Rock Center:

```{r}
rock_map <- ggplot(data = tree_states, mapping = aes(x = long, y = lat, group = group)) + 
  geom_polygon(color = "black", fill = "white") +
  geom_point(aes(y=40.759358, x=-73.978502), color = "darkgreen", size = 5) +
  coord_map("conic", lat0 = 40) +
  theme_bw()
rock_map
```

Map states with Rock Center and locations of where Rockefeller trees came from:

```{r}
rock_map2 <- ggplot(data = tree_states, 
                      mapping = aes(x = long, y = lat, group = group)) + 
  geom_polygon(color = "black", fill = "white") +
  geom_point(data = tree, aes(x = OriginLong, y = OriginLat), 
             color = "red", size = 2, inherit.aes = F) +
  geom_point(aes(y=40.759358, x=-73.978502), 
             color = "darkgreen", size = 5) +
  coord_map("conic", lat0 = 40) +
  theme_bw()
rock_map2
```


Calculate distance to Rock Center:

```{r}
tree <- tree %>% 
  mutate(LatRock = 40.759358,
         LongRock = -73.978502,
         dist_to_rock_m = distHaversine(cbind(OriginLong, OriginLat), 
                              cbind(LongRock, LatRock)),
         dist_to_rock_miles = dist_to_rock_m * 0.000621371)
```

Make a violin plot showing the Distance to Rockefeller Center (miles):

```{r}
rock_box <- ggplot(data = tree, aes(x = dist_to_rock_miles, y = 1)) + 
  geom_violin() +
  theme_bw()+
  xlab("Distance to Rockefeller Center (miles)")+
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank())
rock_box
```

Filter data to get close/far trees:

```{r}
tree_loc <- tree %>% 
  filter(OriginCity %in% c("Wayne, NJ", "Richfield, OH", "Rockefeller Center") & Year %in% c(1998, 2005, NA)) 
  
```

Map closest and farthest trees and label them:

```{r}
county_map2 <- ggplot(data = tree_states, 
                      mapping = aes(x = long, y = lat, group = group)) + 
  geom_polygon(color = "black", fill = "white") +
  geom_point(data = tree_loc, aes(x = OriginLong, y = OriginLat), 
             color = "red", size = 2, inherit.aes = F) +
    geom_point(aes(y=40.759358, x=-73.978502), 
             color = "darkgreen", size = 5) +
  geom_text_repel(data = tree_loc, aes(x = OriginLong, y = OriginLat, label = OriginCity), 
        fontface = "bold", nudge_y = -1, nudge_x = -2,
        inherit.aes = F) +
    coord_map("conic", lat0 = 40) +
  theme_bw()
county_map2
```

